---
title: "Assignment Course Project 1"
author: "Richard Yin"
date: "2016¦~2¤ë19¤é"
output: 
  html_document: 
    keep_md: yes
---


```{r, echo=TRUE}
plot(cars)
```
## loading lattice library
```{r, echo=TRUE}
library(lattice)

## loading dataset from local drive
```{r, echo=TRUE}
MonitoringData <- read.csv("activity.csv")
```

## checking the dimension and first few row of dataset
```{r, echo=TRUE}
dim(MonitoringData)
```
# [1] 17568     3

```{r, echo=TRUE}
head(MonitoringData)
```
# steps       date interval
# 1    NA 2012-10-01        0
# 2    NA 2012-10-01        5
# 3    NA 2012-10-01       10
# 4    NA 2012-10-01       15
# 5    NA 2012-10-01       20
# 6    NA 2012-10-01       25

## calculate the total number of steps taken per day 
```{r, echo=TRUE}
SummaryStepsDataByDate <- aggregate(MonitoringData$steps, 
                                    by = list(MonitoringData$date), 
                                    FUN = sum, 
                                    na.rm = TRUE)
```

## rename the column in Summary dataset
```{r, echo=TRUE}
names(SummaryStepsDataByDate) <- c("date", "total")
```

## display the first few row of total number of steps taken per day
```{r, echo=TRUE}
head(SummaryStepsDataByDate)
```
# date total
# 1 2012-10-01     0
# 2 2012-10-02   126
# 3 2012-10-03 11352
# 4 2012-10-04 12116
# 5 2012-10-05 13294
# 6 2012-10-06 15420

## plot the histogram of the total number of steps taken per day
```{r, echo=TRUE}
hist(SummaryStepsDataByDate$total,
     breaks = seq(from = 0, to = 25000, by = 1000),
     xlab = "Total number of steps",
     ylim = c(0, 12),
     main = "Histogram of the total number of steps taken per day\n NA been removed"
     )
```

## calculate the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
TotalNumberofMean <- mean(SummaryStepsDataByDate$total)
TotalNumberofMedian <- median(SummaryStepsDataByDate$total)
```

## display the mean and median of the total number of stps taken per day on screen
```{r, echo=TRUE}
TotalNumberofMean
```
# [1] 9354.23

```{r, echo=TRUE}
TotalNumberofMedian
```
# [1] 10395

## calculate the average number of steps taken across all day 
```{r, echo=TRUE}
SummaryStepsDataByInterval <- aggregate(MonitoringData$steps,
                                        by = list(MonitoringData$interval),
                                        FUN = mean, 
                                        na.rm = TRUE)
```

## rename the column in Summary dataset
```{r, echo=TRUE}
names(SummaryStepsDataByInterval) <- c("interval", "mean")
```

## display the first few row of total number of steps across all day
```{r, echo=TRUE}
head(SummaryStepsDataByInterval)
```
#   interval      mean
# 1        0 1.7169811
# 2        5 0.3396226
# 3       10 0.1320755
# 4       15 0.1509434
# 5       20 0.0754717
# 6       25 2.0943396

## plot the time series plot which steps taken across all days
```{r, echo=TRUE}
plot(SummaryStepsDataByInterval$interval,
     SummaryStepsDataByInterval$mean,
     type = "l",
     xlab = "Interval in minutes",
     ylab = "Average number of steps",
     main = "Average number of steps per interval\n NA been removed")
```

## find the maximun interval one
```{r, echo=TRUE}
MaximumLocation <- which(SummaryStepsDataByInterval$mean == max(SummaryStepsDataByInterval$mean))
MaximumOne <- SummaryStepsDataByInterval[MaximumLocation, 1]
```

## display the maximun one
```{r, echo=TRUE}
MaximumOne
```
# [1] 835

## count the NA in monitoring data dataset
```{r, echo=TRUE}
CountNA <- sum(is.na(MonitoringData$steps))
```

## display the NA counting result
```{r, echo=TRUE}
CountNA
```
# [1] 2304

## find the NA location
```{r, echo=TRUE}
NALocation <- which(is.na(MonitoringData$steps))
```

## filling in all of the missing value in the dataset by using the mean of the steps value
```{r, echo=TRUE}
ReplicateData <- rep(mean(MonitoringData$steps, na.rm = TRUE),
                     times = length(NALocation))
MonitoringData[NALocation, 1] <- ReplicateData
```

## display the first few row of filling in missing value in the dataset
```{r, echo=TRUE}
head(MonitoringData)
```
# steps       date interval
# 1 37.3826 2012-10-01        0
# 2 37.3826 2012-10-01        5
# 3 37.3826 2012-10-01       10
# 4 37.3826 2012-10-01       15
# 5 37.3826 2012-10-01       20
# 6 37.3826 2012-10-01       25

## calculate the total number of steps taken per day after NA filled up
```{r, echo=TRUE}
SummaryStepsDataByDateNONA<- aggregate(MonitoringData$steps, 
                                    by = list(MonitoringData$date), 
                                    FUN = sum, 
                                    na.rm = TRUE)
```

## rename the column in Summary dataset after NA filled up
```{r, echo=TRUE}
names(SummaryStepsDataByDateNONA) <- c("date", "total")
```

## display the first few row of total number of steps taken per day after NA filled up
```{r, echo=TRUE}
head(SummaryStepsDataByDateNONA)
```
# date    total
# 1 2012-10-01 10766.19
# 2 2012-10-02   126.00
# 3 2012-10-03 11352.00
# 4 2012-10-04 12116.00
# 5 2012-10-05 13294.00
# 6 2012-10-06 15420.00

## plot the histogram of the total number of steps taken per day after NA filled up
```{r, echo=TRUE}
hist(SummaryStepsDataByDateNONA$total,
     breaks = seq(from = 0, to = 25000, by = 1000),
     xlab = "Total number of steps",
     ylim = c(0, 20),
     main = "Histogram of the total number of steps taken per day\n NA been filled up"
)
```

## calculate the mean and median of the total number of steps taken per day after NA filled up
```{r, echo=TRUE}
TotalNumberofMeanNONA <- mean(SummaryStepsDataByDateNONA$total)
TotalNumberofMedianNONA <- median(SummaryStepsDataByDateNONA$total)
```

## display the mean and median of the total number of stps taken per day on screen after NA filled up
```{r, echo=TRUE}
TotalNumberofMeanNONA
```
# [1] 10766.19

##  after imputing missing data in the dataset, mean and median value will change to biger one
```{r, echo=TRUE}
TotalNumberofMedianNONA
```
# [1] 10766.19

## transform date in dataset to the actual date format
```{r, echo=TRUE}
MonitoringData$date <- as.POSIXct(MonitoringData$date, format = "%Y-%m-%d")
```

## add new column and transfer date to weekdays
```{r, echo=TRUE}
MonitoringData <- data.frame(date = MonitoringData$date,
                             steps = MonitoringData$steps,
                             interval = MonitoringData$interval,
                             weekday = weekdays(MonitoringData$date)
                             )
```

## display first few row of dataset
```{r, echo=TRUE}
head(MonitoringData)
```
# date   steps interval weekday
# 1 2012-10-01 37.3826        0  Monday
# 2 2012-10-01 37.3826        5  Monday
# 3 2012-10-01 37.3826       10  Monday
# 4 2012-10-01 37.3826       15  Monday
# 5 2012-10-01 37.3826       20  Monday
# 6 2012-10-01 37.3826       25  Monday

## add another new column for date type
```{r, echo=TRUE}
MonitoringData <- cbind(MonitoringData, 
                        datetype = ifelse(MonitoringData$weekday == "Saturday" 
                                          | MonitoringData$weekday == "Sunday",
                                          "Weekend", "Weekday"))
```

## display first few row of dataset
```{r, echo=TRUE}
head(MonitoringData)
```
# date   steps interval weekday datetype
# 1 2012-10-01 37.3826        0  Monday  Weekday
# 2 2012-10-01 37.3826        5  Monday  Weekday
# 3 2012-10-01 37.3826       10  Monday  Weekday
# 4 2012-10-01 37.3826       15  Monday  Weekday
# 5 2012-10-01 37.3826       20  Monday  Weekday
# 6 2012-10-01 37.3826       25  Monday  Weekday

## calculate the average number of steps taken across all weekday or weekend days
```{r, echo=TRUE}
SummaryStepsDataByDateType <- aggregate(MonitoringData$steps,
                                        by = list(MonitoringData$interval,
                                                  MonitoringData$weekday,
                                                  MonitoringData$datetype),
                                        mean)
```

## rename the column in Summary dataset
```{r, echo=TRUE}
names(SummaryStepsDataByDateType) <- c("interval", "weekday", "datetype", "mean")
```

## display first few row of dataset
```{r, echo=TRUE}
head(SummaryStepsDataByDateType)
```
# interval weekday datetype     mean
# 1        0  Friday  Weekday 8.307244
# 2        5  Friday  Weekday 8.307244
# 3       10  Friday  Weekday 8.307244
# 4       15  Friday  Weekday 8.307244
# 5       20  Friday  Weekday 8.307244
# 6       25  Friday  Weekday 8.307244

## plot the time series plot which steps taken across all weekday or weekend days
```{r, echo=TRUE}
xyplot(mean ~ interval | datetype, 
       SummaryStepsDataByDateType,
       type = "l", 
       layout = c(1, 2),
       xlab = "Interval",
       ylab = "Number of steps taken across all\n weekday or weekend"
       )
```


